### 2021-11-24 article.tex
(Ab)Using Implicits

\subsection{A Lightweight Form of Dependent Typing}
Match types enable a lightweight form of dependent typing, since a term-level pattern matching expression can be typed accordingly at the type level as a match type.
Consider the following function definition:
%
\begin{lstlisting}
def elem[X](x: X): Elem[X] = x match {
  case x: String => x.charAt(0)
  case x: List[t] => elem(x.head)
  case x: Any => x
}
\end{lstlisting}
%
This definition is well-typed because the match expression in \dotty{elem}'s body has the exact same scrutinee and pattern types as \dotty{Elem[X]} (the function's return type).

@@

Haskell's type families allow programmers to define type-level functions using pattern matching~\citep{chakravarty2005associated, schrijvers2008open}.
By default, type families are open, which means that a definition can spread across multiple files and compilations.
This flexibility induces a substantial restriction on type family definitions: patterns must not overlap.
One benefit of this restriction is that it prevents any ambiguity in the reduction of type families (patterns are pairwise disjoint), which is required given the distributed nature of definitions.
Open type families are well-suited to be used in conjunction with type classes, since both constructs have open-ended definitions with non-overlapping constraints.

>>>>

\subsection{A Lightweight Form of Dependent Typing}
Match types enable a lightweight form of dependent typing, since term-level pattern matching expressions can be typed accordingly at the type level as a match types.
Consider the following function definition:
%
\begin{lstlisting}
def elem[X](x: X): Elem[X] = x match {
  case x: String => x.charAt(0)
  case x: List[t] => elem(x.head)
  case x: Any => x
}
\end{lstlisting}
%
This definition is well-typed because the match expression in \dotty{elem}'s body has the exact same scrutinee and pattern types as \dotty{Elem[X]} (the function's return type).

@@

Haskell's type families~\citep{chakravarty2005associated, schrijvers2008open} allow programmers to define type-level functions using pattern matching.
By default, type families are open, which means that a definition can spread across multiple files and compilations.
This flexibility induces a substantial restriction on type family definitions: patterns must not overlap.
One benefit of this restriction is that it prevents any ambiguity in the reduction of type families (patterns are pairwise disjoint), which is required given the distributed nature of definitions.
Open type families are well-suited to be used in conjunction with type classes, since both constructs have open-ended definitions with non-overlapping constraints.
